---
title: "Florida Man meets COVID-19"
output: 
  revealjs::revealjs_presentation:
    theme: serif
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# We've got some data!

YAY! 

## Main Research Question/Variable
 Priorities for the economy versus getting the virus under control.

 >- *Getting the virus under control is essential before we open up the economy*

 >- *We should open the economy so that people can get back to their jobs and normal lives*

 >- *[VOLUNTEERED] Both/somewhere in between*

## Hypotheses & Research Questions

>- Will **economic anxieties** or **health anxieties** drive people towards one of these preferences?

>- Will **_racial identity considerations_** drive these preferences?


# Exploring the Data
## Needed R packages
```{r packages, eval=T, echo=T, message=FALSE, warning=FALSE, include=TRUE}
library(readstata13)
library(tidyverse)
library(nnet)
library(stargazer)
```


## Explore the DV

```{r echo=FALSE}
setwd("D:/GitHub/Research-Projects/FLCovid_2020/")
data_raw <- read.dta13("data/BEBR_final.dta")


##Helper Recode Functions
#NAing all negative values
to_na <- function(x){
  ifelse(x < 0, NA, x )
  
}

#0/1 dummies.
make_dum <- function(x){
  ifelse(x == 2, 0 , x)
}

data_cleaned <- data_raw %>%
  #Sets all values less than zero to NA
  mutate(across(.cols = everything(), to_na),
         #Dummies Sex(1=Male), Voter(1=Registered), and Hispanic(1=Hispanic)
         across(.cols = c(ISEX, VOTER, HISPAN), make_dum),
         #Racial Dummies (Other and mixed together; intended reference category)
         white = ifelse(data_raw$RRACE==1 & HISPAN !=1,1,0),
         black = ifelse(data_raw$RRACE==2 & HISPAN !=1,1,0),
         asian = ifelse(data_raw$RRACE==3 & HISPAN !=1,1,0),
         native = ifelse(data_raw$RRACE==4 & HISPAN !=1,1,0),
         other = ifelse(data_raw$RRACE>4& HISPAN !=1,1,0),
         #PID dummies (Other/nopref reference)
         REP = ifelse(data_raw$PID==2,1,0),
         DEM = ifelse(data_raw$PID==1,1,0)) %>%
  #Renaming our variables
  rename(HealthAnxious = MCD1,
         FinanceAnxious = MCD2,
         UnderControl = MCD3,
         RaceImport = MCD4)




#Take a look at distribution of answers on the DV of interest

#Data labels
whatdo<-c("1" ="Getting the virus\nunder control is\n essential before\nwe open up\nthe economy",
          "2" = "We should open\nthe economy so that\n people can get\nback to their\n jobs and\nnormal lives",
          "3" = "Both/somewhere\nin between \n[Volunteered]")

(p1<-ggplot()+geom_bar(data = filter(data_cleaned, !is.na(UnderControl)),
                       aes(x=as.factor(UnderControl)), fill = "#84CEFF" )+theme_minimal()+
  scale_x_discrete(labels = whatdo, name = "")+
    labs(y = "Number of\nRespondents", 
         title = " Floridians' Thoughts on Whether\nto Prioritize Virus vs. Economy")+
    theme(axis.title.y = element_text(hjust = 1, angle = 0, size = 8),
          plot.title = element_text(hjust = .5)) )
```

## Explore the IVs

```{r,results = 'asis'}
stargazer(data_cleaned[c("white","HISPAN","EDUCAT","INCOM2", "black","asian","native","other","REP","DEM","HealthAnxious","FinanceAnxious","UnderControl","RaceImport")], type = "html", digits = 2)
```


## Here be Coding

##Import the data
```{r import, eval=F, echo=T, include=TRUE}
setwd("D:/GitHub/Research-Projects/FLCovid_2020/")
data_raw <- read.dta13("data/BEBR_final.dta")

```

## Cleaning 
```{r clean, eval=F, echo=T, include=TRUE}
to_na <- function(x){
  ifelse(x < 0, NA, x )
  
}

#0/1 dummies.
make_dum <- function(x){
  ifelse(x == 2, 0 , x)
}

data_cleaned <- data_raw %>%
  #Sets all values less than zero to NA
  mutate(across(.cols = everything(), to_na),
         #Dummies Sex(1=Male), Voter(1=Registered), and Hispanic(1=Hispanic)
         across(.cols = c(ISEX, VOTER, HISPAN), make_dum),
         #Racial Dummies (Other and mixed together; intended reference category)
         white = ifelse(data_raw$RRACE==1 & HISPAN !=1,1,0),
         black = ifelse(data_raw$RRACE==2 & HISPAN !=1,1,0),
         asian = ifelse(data_raw$RRACE==3 & HISPAN !=1,1,0),
         native = ifelse(data_raw$RRACE==4 & HISPAN !=1,1,0),
         other = ifelse(data_raw$RRACE>4& HISPAN !=1,1,0),
         #PID dummies (Other/nopref reference)
         REP = ifelse(data_cleaned$PID==2,1,0),
         DEM = ifelse(data_cleaned$PID==1,1,0)) %>%
  #Renaming our variables
  rename(HealthAnxious = MCD1,
         FinanceAnxious = MCD2,
         UnderControl = MCD3,
         RaceImport = MCD4)


```


# Modeling

>- If only a few people picked both then we could probably just nix them and do logit.

>- So many people are doing it though, and these are categorical in nature, so we're gonna need a model that accomodates that.

>- Enter: Multinomial Logit

## But first... -.-

If you don't remove the NAs, then it's considered a "factor" 

```{r pruning, echo=T, eval=F, include=TRUE}
data_pruned <- data_cleaned[!is.na(data_cleaned$UnderControl),]
data_pruned$UnderControl<-as.factor(data_pruned$UnderControl)
data_pruned$UnderControl<-relevel(data_pruned$UnderControl,ref = "3")
```

## Four models (Different Bases & Interactions)

##Non-interaction (Open up as Base)
```{r include=FALSE}
data_pruned <- data_cleaned[!is.na(data_cleaned$UnderControl),]

data_pruned$UnderControl<-as.factor(data_pruned$UnderControl)

data_pruned$UnderControl<-relevel(data_pruned$UnderControl,ref = "3")


#Setting the Multinomial Model
model1_3<- multinom(UnderControl ~ FinanceAnxious + HealthAnxious + 
                    INCOM2 + EDUCAT + AGE + HISPAN +
                    REP + DEM + white + black + asian ,data = data_pruned)




#Base as 2 now

data_pruned$UnderControl<-relevel(data_pruned$UnderControl,ref = "2")

#Setting the Multinomial Model
model1_2<- multinom(UnderControl ~ FinanceAnxious + HealthAnxious + 
                    INCOM2 + EDUCAT + AGE + HISPAN +
                    REP+ DEM+ white + black + asian ,data = data_pruned)

```

```{r, results='asis'}
stargazer(model1_2, column.sep.width = "1pt", font.size = "tiny" ,single.row = T, type = "html", style = "ajps", column.labels = c("Virus Under Control", "Both"),dep.var.labels.include = FALSE,
          model.numbers          = FALSE)
```

## Non-interaction (Both as Base)
```{r,  results='asis'}
stargazer(model1_3, column.sep.width = "1pt", font.size = "tiny" ,single.row = T, type = "html", style = "ajps", column.labels = c("Virus Under Control", "Open Economy"),dep.var.labels.include = FALSE,
          model.numbers          = FALSE)
```

